Characterizing Signal Loss in the 21\,cm Reionization Power Spectrum: \\A Revised Study of PAPER-64

### 2018-08-30 PAPER.tex
read-through edits up until the end

In our toy model, we specifically choose toy models where the data covariance is diagonal, as indeed we expect the EoR signal to be. We assume we have $N$ data points $\Delta_i$ which are the sum of a desired signal $\sigma_i$ and an undesired contaminant $\upsilon_i$
\begin{equation}
\Delta_i = \sigma_i + \upsilon_i
\end{equation}
with 
\begin{equation}
\langle \sigma_i \rangle = 0; \; \langle \sigma^2_i \rangle = s; \; {\rm and}~\langle \bm{\sigma \sigma}^T \rangle = s \mathbf{I}_{N \times N} \equiv \mathbf{S},
\end{equation}
where we wish to estimate $s$.  The contaminant in this first case has a similar structure (as the EoR) for its covariance, and is assumed uncorrelated with the signal
\begin{equation}
\label{eq:IdealToyModelCovariance}
\langle \upsilon_i \rangle = 0; \; 
\langle \upsilon^2_i \rangle = u; \; 
\langle \bm{\upsilon \upsilon}^T \rangle = u \mathbf{I}_{N \times N} \equiv \mathbf{U}; {\rm and}~ 
\langle \sigma_i \upsilon_j \rangle = 0.
\end{equation}
With the covariance matrix given by $\C = \mathbf{S} + \mathbf{U}$, the estimator for $s$ using only the quadratic part of Equation \ref{eq:OQE} is
\begin{equation}
\label{eq:IdealToyModelEstimator}
\hat{s} = \frac{ \bm{\Delta^T \Delta}}{N} 
\end{equation}
and its expectation is
\begin{equation}
\langle \hat{s} \rangle = s + u.
\end{equation}
Thus, {\it when the covariance structure of the contaminant is identical to the signal} ($\PDeriv{\mathbf{S}}{s} = \PDeriv{\mathbf{U}}{u} =  \PDeriv{\C}{s}$), the information available to the quadratic portion of the estimator to distinguish between the two is degenerate, and knowledge only of $\C$ and $\PDeriv{\C}{s}$ is inadequate.  In order to obtain an unbiased estimate of $s$, one must also use knowledge of $\mathbf{U}$.  Indeed, computing the linear bias from Equation \eqref{eq:OQELinear}, one finds $b = u$.

@@

\noindent where $\sigma$ is the standard deviation of $\widehat{P}_{\rm out}$ and $\bar y$ is the mean of $\widehat{P}_{\rm out}$, and they are both functions of $P_{\rm in}$. Using Equations \eqref{eq:prob} and \eqref{eq:logprob}, the quantity inside the expectation value of Equation \eqref{eq:jeffreys} becomes:

%\begin{equation}
%\label{eq:logprob_full}
%\mathcal{L} = -\, log \, \sqrt{2\pi} - log\, \sigma - \frac{1}{2}\Big(\frac{y-\bar y}{\sigma}\Big)^{2}. 
%\end{equation}

%Next, the quantity inside the expectation value of Equation \eqref{eq:jeffreys} can be computed by taking the derivative of %Equation \eqref{eq:logprob_full} with respect to $P_{\rm in}$, or $x$:

\begin{eqnarray}
\Big(\frac{\partial \mathcal{L}}{\partial x} \Big)^{2} &=& \frac{1}{\sigma^{2}}\Big(\frac{\partial \sigma}{\partial x}\Big)^{2} -  \Big(\frac{2(y-\bar y)}{\sigma^{3}}\Big)\frac{\partial \sigma}{\partial x}\frac{\partial \bar y}{\partial x} - \Big(\frac{2(y-\bar y)^{2}}{\sigma^{4}}\Big)\Big(\frac{\partial \sigma}{\partial x}\Big)^{2} \nonumber \\
&+& \Big(\frac{(y-\bar y)^{2}}{\sigma^{4}}\Big)\Big(\frac{\partial \bar y}{\partial x}\Big)^{2} + \Big(\frac{2(y-\bar y)^{3}}{\sigma^{5}}\Big)\frac{\partial \sigma}{\partial x}\frac{\partial \bar y}{\partial x} + \Big(\frac{(y-\bar y)^{4}}{\sigma^{6}}\Big)\Big(\frac{\partial \sigma}{\partial x}\Big)^{2}.
\end{eqnarray}

Taking the expectation value then removes all terms with odd powers of $(y - \bar y)$ because those Gaussian moments evaluate to zero. Additionally, the second moment can be simplified since $\langle (y - \bar y)^{2} \rangle = \sigma^{2}$ and the fourth moment can be simplified since $\langle (y - \bar y)^{4} \rangle = 3\sigma^{4}$. Finally, after some additional simplification the Jeffreys prior becomes:

@@

When we simulate our full injection framework as in Section \ref{sec:Practice}, we sample 50 $P_{\rm in}$ values that range from $\sim$ $10^{5}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$ to $\sim$ 10$^{11}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, and we note that the prior is set to zero outside those regions. For the injections that we do sample, we can simply fit analytic functions to the mean and standard deviations of $\widehat{P}_{\rm out}$ ($\bar y$ and $\sigma$) as functions of $P_{\rm in}$. An example of the typical shape of these functions for the PAPER-64 analysis is shown in Figure \ref{fig:jeffreys1}, though in practice we fit solutions for every $k$-value and simulation independently.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/jeffrey_fits.pdf}
	\caption{An illustrative example (for the PAPER-64 analysis using uniform weighting and $k=0.393$\,$h$ Mpc$^{-1}$) of how the mean of $P_{\rm out}$ (left) and standard deviation of $P_{\rm out}$ (right) behave as a function of $P_{\rm in}$. Polynomials are fit to each (red) to describe how $\bar y$ and $\sigma$ evolve with $x$ (injection level), respectively, for the computation of the Jeffreys prior as defined in Equation \eqref{eq:jeffreys_final}. The polynomial fits for this example are $y = (-5.1 \times 10^{-15})x^{2} + x + (1.5 \times 10^{7})$ and $y = (5.0 \times 10^{-13})x^{2} + 0.2 x + 10^{7}$ for $\bar y$ and $\sigma$, respectively.}
	\label{fig:jeffreys1}
\end{figure}

We also show the typical shape of the Jeffreys prior used in our analysis in Figure \ref{fig:jeffreys2}, as computed by Equation \eqref{eq:jeffreys_final}. Most noticeably, it is not constant with $P_{\rm in}$, meaning a uniform prior, which is often used for simplicity, is informative in our application. Therefore, due to its objective nature we choose to use a Jeffreys prior in our analysis, multiplying our likelihood functions by Equation \eqref{eq:jeffreys_final} before computing posterior distributions.

>>>>

We specifically choose toy models where the data covariance is diagonal, as indeed we expect the EoR signal to be. We assume we have $N$ data points $\Delta_i$ which are the sum of a desired signal $\sigma_i$ and an undesired contaminant $\upsilon_i$
\begin{equation}
\Delta_i = \sigma_i + \upsilon_i
\end{equation}
with 
\begin{equation}
\langle \sigma_i \rangle = 0; \; \langle \sigma^2_i \rangle = s; \; {\rm and}~\langle \bm{\sigma \sigma}^T \rangle = s \mathbf{I}_{N \times N} \equiv \mathbf{S},
\end{equation}
where we wish to estimate $s$.  The contaminant in this first case has a similar structure (as the EoR) for its covariance, and is assumed uncorrelated with the signal
\begin{equation}
\label{eq:IdealToyModelCovariance}
\langle \upsilon_i \rangle = 0; \; 
\langle \upsilon^2_i \rangle = u; \; 
\langle \bm{\upsilon \upsilon}^T \rangle = u \mathbf{I}_{N \times N} \equiv \mathbf{U}; {\rm and}~ 
\langle \sigma_i \upsilon_j \rangle = 0.
\end{equation}
With the covariance matrix given by $\C = \mathbf{S} + \mathbf{U}$, the estimator for $s$ using only the quadratic part of Equation \ref{eq:OQE} is
\begin{equation}
\label{eq:IdealToyModelEstimator}
\hat{s} = \frac{ \bm{\Delta^T \Delta}}{N} 
\end{equation}
and its expectation is
\begin{equation}
\langle \hat{s} \rangle = s + u.
\end{equation}
Thus, {\it when the covariance structure of the contaminant is identical to the signal} ($\PDeriv{\mathbf{S}}{s} = \PDeriv{\mathbf{U}}{u} =  \PDeriv{\C}{s}$), the information available to the quadratic portion of the estimator to distinguish between the two is degenerate, and knowledge only of $\C$ and $\PDeriv{\C}{s}$ is inadequate.  In order to obtain an unbiased estimate of $s$, one must also use knowledge of $\mathbf{U}$.  Indeed, computing the linear bias from Equation \eqref{eq:OQELinear}, one finds $b = u$.

@@

\noindent where $\sigma$ is the standard deviation of $\widehat{P}_{\rm out}$ and $\bar y$ is the mean of $\widehat{P}_{\rm out}$, and they are both functions of $P_{\rm in}$. Combining Equations \eqref{eq:jeffreys}, \eqref{eq:logprob}, and \eqref{eq:prob} the Jeffreys prior simplifies to become:

%\begin{eqnarray}
%\Big(\frac{\partial \mathcal{L}}{\partial x} \Big)^{2} &=& \frac{1}{\sigma^{2}}\Big(\frac{\partial \sigma}{\partial x}\Big)^{2} -  \Big(\frac{2(y-\bar y)}{\sigma^{3}}\Big)\frac{\partial \sigma}{\partial x}\frac{\partial \bar y}{\partial x} - \Big(\frac{2(y-\bar y)^{2}}{\sigma^{4}}\Big)\Big(\frac{\partial \sigma}{\partial x}\Big)^{2} \nonumber \\
%&+& \Big(\frac{(y-\bar y)^{2}}{\sigma^{4}}\Big)\Big(\frac{\partial \bar y}{\partial x}\Big)^{2} + \Big(\frac{2(y-\bar y)^{3}}{\sigma^{5}}\Big)\frac{\partial \sigma}{\partial x}\frac{\partial \bar y}{\partial x} + \Big(\frac{(y-\bar y)^{4}}{\sigma^{6}}\Big)\Big(\frac{\partial \sigma}{\partial x}\Big)^{2}.
%\end{eqnarray}

%Taking the expectation value then removes all terms with odd powers of $(y - \bar y)$ because those Gaussian moments evaluate to zero. Additionally, the second moment can be simplified since $\langle (y - \bar y)^{2} \rangle = \sigma^{2}$ and the fourth moment can be simplified since $\langle (y - \bar y)^{4} \rangle = 3\sigma^{4}$. Finally, after some additional simplification the Jeffreys prior becomes:

@@

When we simulate our full injection framework as in Section \ref{sec:Practice}, we sample 50 $P_{\rm in}$ values that range from $\sim$ $10^{5}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$ to $\sim$ 10$^{11}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, and we note that the prior is set to zero outside those regions. For the injections that we do sample, we can simply fit analytic functions to the mean and standard deviations of $\widehat{P}_{\rm out}$ ($\bar y$ and $\sigma$) as functions of $P_{\rm in}$. %An example of the typical shape of these functions for the PAPER-64 analysis is shown in Figure \ref{fig:jeffreys1}, though in practice we fit solutions for every $k$-value and simulation independently.

%\begin{figure}
%	\centering
%	\includegraphics[width=\columnwidth]{plots/jeffrey_fits.pdf}
%	\caption{An illustrative example (for the PAPER-64 analysis using uniform weighting and $k=0.393$\,$h$ Mpc$^{-1}$) of how the mean of $P_{\rm out}$ (left) and standard deviation of $P_{\rm out}$ (right) behave as a function of $P_{\rm in}$. Polynomials are fit to each (red) to describe how $\bar y$ and $\sigma$ evolve with $x$ (injection level), respectively, for the computation of the Jeffreys prior as defined in Equation \eqref{eq:jeffreys_final}. The polynomial fits for this example are $y = (-5.1 \times 10^{-15})x^{2} + x + (1.5 \times 10^{7})$ and $y = (5.0 \times 10^{-13})x^{2} + 0.2 x + 10^{7}$ for $\bar y$ and $\sigma$, respectively.}
%	\label{fig:jeffreys1}
%\end{figure}

We show the typical shape of the Jeffreys prior used in our analysis in Figure \ref{fig:jeffreys2}, as computed by Equation \eqref{eq:jeffreys_final}. Most noticeably, it is not constant with $P_{\rm in}$, meaning a uniform prior, which is often used for simplicity, is informative in our application. Therefore, due to its objective nature we choose to use a Jeffreys prior in our analysis, multiplying our likelihood functions by Equation \eqref{eq:jeffreys_final} before computing posterior distributions.
